---
title: "HW10"
author: "Nihar Harkare"
date: "2024-04-23"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    toc_float: yes
---

[https://github.com/hinarh/HW10sds315](https://github.com/hinarh/HW10sds315)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(ggplot2)
```

# Problem 1

```{r echo=FALSE}
redline = read_csv("redlining.csv", show_col_types = FALSE)

lm_model = lm(policies ~ minority + fire + age + income, data = redline)

summary(lm_model)

min_conf = confint(lm_model, "minority", level = 0.95)
```
1) Question:

Is there an association between the number of FAIR policies and the racial/ethic composition of a ZIP code?

2) Approach:

To answer this question, I used a linear regression model, andy including independent variables in the model, we can control for their effects and better understand the specific association between FAIR policies and racial/ethnic composition of a ZIP code.

3) Results:

Minority (racial/ethnic composition): The coefficient for minority is 0.008, and the associated p-value is 0.006.

Fire Incidents: The coefficient for fire incidents is 0.022 with a p-value of 0.018.

Age of Housing Units: The coefficient for age of housing units is 0.006 with a p-value of 0.143.

Median Family Income: The coefficient for median family income is -0.016 with a p-value of 0.679.

4) Conclusion:

We can see, since the p-value is very low, the coefficient is significant, meaning that as the number of fair policies increase, the percentage of minorities increases by very little, from `r min_conf[0]` to `r min_conf[1]` at 95% confidence.

# Problem 2

## Part A

```{r echo=FALSE}
groceries = read_csv("groceries.csv", show_col_types = FALSE)

average_price_per_store = groceries %>%
  group_by(Store) %>%
  summarize(Average_Price = mean(Price))

ggplot(average_price_per_store, aes(x = Average_Price, y = reorder(Store, Average_Price))) +
  geom_bar(stat = "identity", fill = "green") +
  labs(title = "Average Price of Products Sold at Different Stores", x = "Average Price (in dollars)", y = "Store") +
  theme(axis.text.y = element_text(angle = 45, hjust = 1))
```

Average price calculated from available products in the dataset.

## Part B

```{r echo=FALSE}
stores_per_product = groceries %>%
  group_by(Product) %>%
  summarize(Num_Stores = n_distinct(Store))

ggplot(stores_per_product, aes(x = Num_Stores, y = reorder(Product, Num_Stores))) +
  geom_bar(stat = "identity", fill = "orange") +
  labs(title = "Number of Stores Selling Each Product", x = "Number of Stores", y = "Product")
```

## Part C

```{r echo=FALSE}
groceries_nona = na.omit(groceries)

convenience_model = lm(Price ~ Product + Type, data = groceries_nona)

conf_interval = confint(convenience_model, "TypeConvenience")

lower_bound = conf_interval[1]
upper_bound = conf_interval[2]

lower_bound = round(lower_bound, 2)
upper_bound = round(upper_bound, 2)
```

Compared with ordinary grocery stores, convenience stores charge between $`r lower_bound`, and $`r upper_bound` more for the same product.

## Part D

```{r echo=FALSE}
store_model = lm(Price ~ Product + Store, data = groceries)

summary(store_model)
```

Looking at the results, Walmart and Kroger seem to charge the lowest prices when comparing the same product, and Whole Foods and Wheatsville Food Co-Op seem to charge the highest prices when comparing the same product.

## Part E

```{r echo=FALSE}
coef_heb = coef(store_model)["StoreHEB"]
coef_cm = coef(store_model)["StoreCentral Market"]
```


Looking at the coefficients, `r coef_heb` for HEB and `r coef_cm` for central market, we can see that Central Market charges a similar amount to HEB for the same product. The difference between these two seem to be the smallest out of any store.

## Part F

```{r echo=FALSE}
groceries = groceries %>%
  mutate(Income10K = Income / 10000)

income_model = lm(Price ~ Product + Income10K, data = groceries)

summary(income_model)

coef_inc10k = coef(income_model)["Income10K"]

sd_income = sd(groceries$Income10K)

change_in_price = coef_inc10k * sd_income
```

Since the coefficient for Income10K is `r coef_inc10k` and the sign is negative, we can assume consumers in poorer ZIP codes seem to pay less for the same product, on average.

A one-standard deviation increase in the income of a ZIP code seems to be associated with
a `r change_in_price` standard-deviation change in the price that consumers in that ZIP code expect to pay for
the same product.